---
title: "Shiny - Part 2"
output: html_document
date: '2022-12-05'
---

#Loading the dataset
```{r pressure, echo=FALSE}
# read in the data
library(tidyverse)
datafile <- "https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv"
health_data <- read_csv(datafile)

# Explore the data
summary(health_data)
str(health_data)
```

#Checking for missing values and Interpolating them
```{r pressure, echo=FALSE}
# After we check the dataframe, we find that there are NA values in the bmi and hypertension columns
# calculate the NA rows
nrow(health_data[is.na(health_data$bmi),])
nrow(health_data[is.na(health_data$hypertension),])
# deal with the missing values
library(imputeTS)
health_data$bmi <- na_interpolation(health_data$bmi)

# use sample to fill NA values in hypertension column, in case generate decimal
health_data$hypertension <- ifelse(is.na(health_data$hypertension),
                                  sample(c(0,1), size = nrow(health_data), replace = TRUE),
                                  health_data$hypertension)
nrow(health_data[is.na(health_data$bmi),])
nrow(health_data[is.na(health_data$hypertension),])
```

#Defining "Expensive" and "Non-Expensive"
```{r}
# According to the distribution of cost, we can define $5779.8 (80% value) as the threshold.
# If the cost is higher than 5779.8, it is expensive; if not, it is non-expensive.
# Source Reference: https://www.ehealthinsurance.com/resources/individual-and-family/how-much-does-individual-health-insurance-cost
```

#Creating Expensive column
```{r}
# create a new column named expensive, if cost > 5779.8, the valuse is 1, otherwise 0
health_data$expensive <- ifelse(health_data$cost > 5779.8, 1, 0)
# check the distribution of the new column
table(health_data$expensive)
# check the proportion of the new column
mean(health_data$expensive)
#View(health_data)
```

#Creating factors of smoker, yearly_physical and exercise
```{r}
health_data$smoker<- as.factor(health_data$smoker)
health_data$yearly_physical<-as.factor(health_data$yearly_physical)
health_data$exercise <- as.factor(health_data$exercise)
```

#Logistic Regression on statistically significant variables only 
```{r}
#Regression Model
mylogit1 <- glm(expensive ~ age + bmi + children + smoker + yearly_physical + exercise + hypertension, data = health_data, family = "binomial")
summary(mylogit1)
```

#SVM model using Logistic Regression training set
```{r}
library(kernlab)
library(caret)
health_data$expensive <- as.factor(health_data$expensive)
training <- createDataPartition(y=health_data$expensive,p=.70,list=FALSE)
train <- health_data[training,] # training set
test <- health_data[-training,] # test set
# create a svm model to predict the expensive column
library(e1071)
svm_model <- train(expensive ~ age + bmi + children + smoker + yearly_physical + exercise + hypertension, data = train, method = "svmRadial", preProcess = c("center", "scale"))
# predict the expensive column for the test data
svm_pred <- predict(svm_model, test)
# create the sensitivity confusion matrix
confusionMatrix(svm_pred,test$expensive)

#Saving Model output
our_model <- svm_model
save(our_model,file = "our_model.rda")
```

#Explorting the file we worked on i.e. "Solution File"
```{r}
write.csv(health_data,"health_db.csv")
```

#Importing the testing file which the user will import without the cost column i.e. "Testing File"
```{r}
#dftest <- read.csv("HMO_data.csv")
dftest1<- health_data[,c(-14)] 
write.csv(dftest1,"dftest1.csv")
```

#Shiny Code for user to input the "Testing File" and uploading the "Solution File" with number of rows
#Shiny outputs the confusion matrix with accuracy and sensitivity
```{r}
library(rsconnect)
library(shiny)

rsconnect::setAccountInfo(name='akashsingh',
			  token='2E278F9AD4E9B9EB0EF3CB00A30B5DC3',
			  secret='HASJ9huVSIiJtPtV3+XVUmAYOjAW/xrPtNS8DRHc')

ui <- fluidPage(
  fileInput("upload",label = "Input File",accept = c(".csv")),
  fileInput("upload_Solution",label = "Solution File",accept = c(".csv")),
  numericInput("n","Number of Rows", value = 5, min = 1, step = 1),
  tableOutput("headforDF"),
  verbatimTextOutput("txt_results",placeholder = TRUE)
)

server <- function(input,output,session) {
  getTestData <- reactive({
    req(input$upload)
    read_csv(input$upload$name)
  })
  
  getSolutionData <- reactive({
    req(input$upload_Solution)
    read_csv(input$upload_Solution$name)
  })
  
  get_n <-reactive({
    req(input$n)
  })
output$txt_results <- renderPrint({
  dataset <- getTestData()
  dataset_solution <- getSolutionData()
  n <-get_n()
  use_model_to_predict(dataset[1:n,],dataset_solution[1:n,])
  })
output$headForDF <- renderTable({
  df <- getTestData()
  head(df,n)
  })
}

library(caret)
library(kernlab)
library(e1071)

use_model_to_predict <- function(df,df_solution){
  load(file="our_model.rda")
  svmPred <- predict(our_model,df,type="raw")
  df_solution$expensive <- as.factor(df_solution$expensive)
  confusionMatrix(svmPred, df_solution$expensive)
}
shinyApp(ui=ui,server=server)
```

